Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-12-21 Paper.Rmd
tasks done : plotting in single pdf for Bivariate data

Question Q1 was answered principally via analysis of the
data using a model selection approach. For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, as in the Introduction, we produced ranks 
$(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; subsequent analysis used these 
*normalized ranks* $u_i$ and $v_i$. Ranks were $1$ for the smallest element in a set. 
Second, we did a test of the independence of the $u_i$ and $v_i$ (implemented in X in the Y package). 
<!--***DAN: X and Y to be flled in, can't remember right now what they are and don't want to stop
to look it up, Shya, you could also fill these in if you get a chance-->
Subsequent model 
selection algorithms are ineffective if data could not be distinguished from independent data 
since most or all bivariate copula families considered include the independent copula in the 
family or as a boundary case. If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit the bivariate copula families listed previously to the ranked data 
$(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`, and 
obtained AIC and BIC values and accompanying model weights, $\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection]. The weight of a model represents the 
relative support of data for the model; weights sum to 1 across models considered. The lower- and 
upper-tail dependence of the best-fitting member of each family was also provided by `BiCopEst`. 
Fourth, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.
Model selection methods will give the *relative* support of several models, but will not,
on their own, indicate whether any of the models selected is an objectively good fit: all 
models considered might poorly describe the data. To test this, and as the final step of
our procedure, we tested the goodness of fit of our AIC-best copula family using a 
bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
two tests, one based on a Cramer-von Mises statistic, and one based on a 
Kolmogorov-Smirnov statistic. For keep computation times reasonable, 
an initial run using $100$ bootstraps was performed, and if
the $p$-value resulting from either test was less than $0.2$, tests were re-run with 
$1000$ boostraps.

@@

\begin{figure}
\begin{center}
\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
\caption{Upper panels shows raw data plots for (A) soil C and N data, (B) $log_{10}$(basal metabolic rate (BMR)) vs. log(body mass) 
for birds, (C) log(BMR) vs. log(body mass) for mammals, and (D) above-ground plant biomass vs. Shannon's index (H) from Cedar Creek. Bottom panels show the corresponding copula plots. \label{fig_biv_multi_raw_cop}}
\end{center}
\end{figure}
<!--***DAN: Shya, we need to produce a single-pdf figure using R, not use latex to
cobble together separate pdfs for each panel. Also, I think we can dispense with the 
plots of the multivariate data. I have created a script that will do both of these things, 
and inserted a call to it in the sup mat Rmd. Right now it only plots dummy, filler data. 
Please look at MakeBivarDataPlot.R (that's the script) and search for
"Shya", and do as it says in those comments, so the actual data are imported into that
script and plotted. I have changed the above so this new figure is plotted instead of the
cobbled together old figures.-->

@@

dt <- tibble(Description = c("1. P-value, independence test",
           "2. Minimum-AIC copula",
           "3. AIC weight of minimum-AIC copula",
           "4. AIC weight of N copula",
           "5. P-value, Cramer-von Mises goodness of fit test",
           "6. P-value, Kolmogorov-Smirnov goodness of fit test",
           "7. Model averaged lower-tail dependence",
           "8. Model averaged upper-tail dependence",
           "9. Model averaged lower- minus upper-tail dependence",
           "10. Asymmetry rank in surrogates")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
knitr::kable(tab_bivar_summary, 
             format="latex",
             caption = "Summary of Q1 results for bivariate datasets. The $p$-values in rows 5-6 and are for the minimum-AIC copula. Model averaging used for rows 7-9 was based on AIC weights. For row 10, the statistic $P_l-P_u$ was computed for real data and for 1000 surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented. \\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar creek data")
             )%>%column_spec(2:5,width="1.4cm")%>%
             column_spec(1, width = "8.5cm")%>%kable_styling(latex_options = "basic",position = "center")
```
<!--***DAN: Shya, it says Pl-Pu above, but I suspect it was P_first slice - P_last slice, right?
Can you please look back at the code, if necessary, and write a comment below this one 
indicating what results are actually presented in the table?-->

>>>>

Question Q1 was answered principally via analysis of the
data using a model selection approach. For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, as in the Introduction, we produced ranks 
$(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; subsequent analysis used these 
*normalized ranks* $u_i$ and $v_i$. Ranks were $1$ for the smallest element in a set. 
Second, we did a test of the independence of the $u_i$ and $v_i$ (implemented in `BiCopIndTest` in the `VineCopula` package). 
<!--***DAN: X and Y to be flled in, can't remember right now what they are and don't want to stop
to look it up, Shya, you could also fill these in if you get a chance
***Shya: Done.-->
Subsequent model 
selection algorithms are ineffective if data could not be distinguished from independent data 
since most or all bivariate copula families considered include the independent copula in the 
family or as a boundary case. If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit the bivariate copula families listed previously to the ranked data 
$(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`, and 
obtained AIC and BIC values and accompanying model weights, $\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection]. The weight of a model represents the 
relative support of data for the model; weights sum to 1 across models considered. The lower- and 
upper-tail dependence of the best-fitting member of each family was also provided by `BiCopEst`. 
Fourth, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.
Model selection methods will give the *relative* support of several models, but will not,
on their own, indicate whether any of the models selected is an objectively good fit: all 
models considered might poorly describe the data. To test this, and as the final step of
our procedure, we tested the goodness of fit of our AIC-best copula family using a 
bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
two tests, one based on a Cramer-von Mises statistic, and one based on a 
Kolmogorov-Smirnov statistic. For keep computation times reasonable, 
an initial run using $100$ bootstraps was performed, and if
the $p$-value resulting from either test was less than $0.2$, tests were re-run with 
$1000$ boostraps.

@@

\begin{figure}
\begin{center}
\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
\caption{Upper panels shows raw data plots for (A) soil C and N data, (B) $log_{10}$(basal metabolic rate (BMR)) vs. log(body mass) 
for birds, (C) log(BMR) vs. log(body mass) for mammals, and (D) above-ground plant biomass vs. Shannon's index (H) from Cedar Creek. Bottom panels show the corresponding copula plots. \label{fig_biv_multi_raw_cop}}
\end{center}
\end{figure}
<!--***DAN: Shya, we need to produce a single-pdf figure using R, not use latex to
cobble together separate pdfs for each panel. Also, I think we can dispense with the 
plots of the multivariate data. I have created a script that will do both of these things, 
and inserted a call to it in the sup mat Rmd. Right now it only plots dummy, filler data. 
Please look at MakeBivarDataPlot.R (that's the script) and search for
"Shya", and do as it says in those comments, so the actual data are imported into that
script and plotted. I have changed the above so this new figure is plotted instead of the
cobbled together old figures.
***Shya: filled in MakeBivDataPlot.R : Done.-->

@@

dt <- tibble(Description = c("1. P-value, independence test",
           "2. Minimum-AIC copula",
           "3. AIC weight of minimum-AIC copula",
           "4. AIC weight of N copula",
           "5. P-value, Cramer-von Mises goodness of fit test",
           "6. P-value, Kolmogorov-Smirnov goodness of fit test",
           "7. Model averaged lower-tail dependence",
           "8. Model averaged upper-tail dependence",
           "9. Model averaged lower- minus upper-tail dependence",
           "10. Asymmetry rank in surrogates")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
knitr::kable(tab_bivar_summary, 
             format="latex",
             caption = "Summary of Q1 results for bivariate datasets. The $p$-values in rows 5-6 and are for the minimum-AIC copula. Model averaging used for rows 7-9 was based on AIC weights. For row 10, the statistic $P_l-P_u$ was computed for real data and for 1000 surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented. \\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar creek data")
             )%>%column_spec(2:5,width="1.4cm")%>%
             column_spec(1, width = "8.5cm")%>%kable_styling(latex_options = "basic",position = "center")
```
<!--***DAN: Shya, it says Pl-Pu above, but I suspect it was P_first slice - P_last slice, right?
Can you please look back at the code, if necessary, and write a comment below this one 
indicating what results are actually presented in the table?
***Shya: Yes, you are right, here Pl-Pu means P_first slice - P_last slice. In table, item 10 : Assymetry rank is shown for Kendall preserving surroagates.
Reference : bivfunctionplot.R file, lines 134,135,161-->

### 2018-12-21 SupportingInformation.Rmd
tasks done : plotting in single pdf for Bivariate data

  sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
}
par(op)
dev.off()
#-----------------------
#raw data plot for year2000
pdf("./Results/Ceder_creek_results/raw_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
dev.off()
#----------------------------------------------------------------------------------------
cop_HB_all_yr<-vector("list",length(yr))
names(cop_HB_all_yr)<-yr

>>>>

  sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
}
par(op)
dev.off()
#-----------------------
#raw data plot for year2000
pdf("./Results/Ceder_creek_results/raw_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
saveRDS(sr,"./Results/Ceder_creek_results/raw_cop_cc_2000.RDS")
dev.off()
#----------------------------------------------------------------------------------------
cop_HB_all_yr<-vector("list",length(yr))
names(cop_HB_all_yr)<-yr
